beautier: BEAUti for R

### 2017-12-22 article.tex
How to build article (for @rsetienne)
EOD

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian}.
BEAUti has a user-friendly graphical user interface to create BEAST2 input files,
but does not allow for the scripted use of BEAST2. '
\verb;beautier; does allow for scripted use of BEAST2 from an R function call.
An additional novelty is that \verb;beautier; allows to specify a fixed crown age.

%\begin{figure}
%  \centering
%  \includegraphics[]{BeautiSiteModel.png}
%  \caption{BEAUti}
%  \label{figure:beauti}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=6cm, semithick]   
  \tikzstyle{every state}=[]
  \node[state] (A) [rectangle] {
    \includegraphics[width=0.2\textwidth]{alignment.png}
    \includegraphics[width=0.2\textwidth]{thought_cloud.png}
  };   
  \node[state] (E) [below of=A, rectangle] {
    \includegraphics[height=0.1\textheight]{xml.png}
  };   
  \node[state] (F) [rectangle, below of=E] {\includegraphics[width=0.3\textwidth]{DensiTreeExample2.jpg}};
  \path (A) edge [anchor = east] node {\includegraphics[height=0.15\textheight]{beautier_logo.png}} (E)
        (A) edge [anchor = west] node {\includegraphics[height=0.15\textheight]{BeautiSiteModel.png}} (E)
        (E) edge [] node {\includegraphics[width=0.2\textwidth]{Beast2LogoRat.png}} (F); 
  \end{tikzpicture}

  \caption{
    Workflow. From an alignment and priors, one creates a BEAST2 XML input file. This
    can be done using beastscript and BEAUti. The XML file created is run by BEAST2
    to create a posterior.
  }
  \label{fig:workflow}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[]
\centering
\begin{tabular}{ | l | l | }
\hline
\textbf{Name} & \textbf{Description} \\
\hline
\verb;create_beast2_input_file; & Creates a BEAST2 input file \\
\hline
\verb;create_gamma_site_model; & Create a gamma site model \\
\verb;create_gtr_site_model; & Create a GTR site model \\
\verb;create_hky_site_model; & Create an HKY site model \\
\verb;create_jc69_site_model; & Create a Jukes-Cantor site model \cite{cantor1969mammalian} \\
\verb;create_tn93_site_model; & Create a TN93 site model \\
\hline
\verb;create_rln_clock_model; & Create a relaxed log-normal clock model \\
\verb;create_strict_clock_model; & Create a strict clock model \\
\hline
\verb;create_bd_tree_prior; & Create a birth-death tree prior \cite{kendall1948generalized} \\
\verb;create_cbs_tree_prior; & Create a coalescent Bayesian skyline tree prior \\
\verb;create_ccp_tree_prior; & Create a coalescent constant-population tree prior \\
\verb;create_cep_tree_prior; & Create a coalescent exponential-population tree prior \\
\verb;create_yule_tree_prior; & Create a Yule tree prior \cite{yule} \\
\hline
\verb;create_beta_distr; & Create a beta distribution \\
\verb;create_exp_distr; & Create an exponential distribution \\
\verb;create_gamma_distr; & Create a gamma distribution \\
\verb;create_inv_gamma_distr; & Create an inverse gamma distribution \\
\verb;create_laplace_distr; & Create a Laplace distribution \\
\verb;create_log_normal_distr; & Create a log-normal distribution \\
\verb;create_normal_distr; & Create a normal distribution \\
\verb;create_one_div_x_distr; & Create a 1/X distribution \\
\verb;create_poisson_distr; & Create a Poisson distribution \\
\verb;create_uniform_distr; & Create a uniform distribution \\
\hline
\end{tabular}
\caption{beautier's functions}
\label{tab:functions}
\end{table}

By creating BEAST2 input files in BEAUti, 
the desired output of \verb;beautier; was determined. 
The functionality of BEAUti is added gradually (and is still growing).

\verb;beautier; has only parts of the functionality of BEAUti, yet
can be extended easily. \verb;beautier;, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

@@

\verb;beautier; has minimal support for parsing and interpreting BEAST2 output files,
for that the \verb;RBeast; \cite{RBeast} package is recommended.

>>>>

BEAST2 \cite{bouckaert2014beast} is a Bayesian phylogenetics tool.
BEAST2 creates a posterior of jointly-estimated phylonies and model
parameters, from a DNA, RNA or amino acid alignment. BEAST2 can be run
from the command line. To run BEAST2 from the command-line, 
an XML configuration file is needed, that contains the alignment and model parameters.

BEAST2 is bundled with BEAUti \cite{drummond2012bayesian}. 
BEAUti is a program to create a BEAST2 XML configuration file, 
with a user-friendly graphical user interface.
BEAUti's purpose is to help new BEAST2 users getting started.
To do so, BEAUti starts with reasonable default settings. To
create a minimal BEAST2 configuration file, BEAUti only needs one alignment file.
BEAUti cannot be used from within shell scripts.

\verb;beautier;, 'BEAUti for R', allows for creating BEAST2 configuration files 
from an R function call. The interface of \verb;beautier; mimics BEAUti. This
familiar interface helps experienced BEAST2 users to create configuration'
files from shell scripts.

BEAUti does not allow yet that an inferred phylogeny's crown age is fixed.
In an emperical context, a phylogeny's crown age is one of the estimated
parameters. For theoretical work, a fixed and known crown age can result
in a cleaner analysis. \verb;beautier; allows to specify a fixed crown age.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=6cm, semithick]   
  \tikzstyle{every state}=[]
  \node[state] (A) [rectangle] {
    \includegraphics[width=0.2\textwidth]{alignment.png}
    \includegraphics[width=0.2\textwidth]{thought_cloud.png}
  };   
  \node[state] (E) [below of=A, rectangle] {
    \includegraphics[height=0.1\textheight]{xml.png}
  };   
  \node[state] (F) [rectangle, below of=E] {\includegraphics[width=0.3\textwidth]{DensiTreeExample2.jpg}};
  \path (A) edge [anchor = east] node {\includegraphics[height=0.15\textheight]{beautier_logo.png}} (E)
        (A) edge [anchor = west] node {\includegraphics[height=0.15\textheight]{BeautiSiteModel.png}} (E)
        (E) edge [] node {\includegraphics[width=0.2\textwidth]{beast_logo.png}} (F); 
  \end{tikzpicture}

  \caption{
    Workflow. From an alignment and priors, one creates a BEAST2 XML input file. This
    can be done using beautier or BEAUti. The created configuration file is run by BEAST2
    to create a posterior.
  }
  \label{fig:workflow}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[]
\centering
\begin{tabular}{ | l | l | }
\hline
\textbf{Name} & \textbf{Description} \\
\hline
\verb;create_beast2_input_file; & Creates a BEAST2 input file \\
\hline
\verb;create_gtr_site_model; & Create a GTR site model \\
\verb;create_hky_site_model; & Create an HKY site model \\
\verb;create_jc69_site_model; & Create a Jukes-Cantor site model \cite{cantor1969mammalian} \\
\verb;create_tn93_site_model; & Create a TN93 site model \\
\hline
\verb;create_rln_clock_model; & Create a relaxed log-normal clock model \\
\verb;create_strict_clock_model; & Create a strict clock model \\
\hline
\verb;create_bd_tree_prior; & Create a birth-death tree prior \cite{kendall1948generalized} \\
\verb;create_cbs_tree_prior; & Create a coalescent Bayesian skyline tree prior \\
\verb;create_ccp_tree_prior; & Create a coalescent constant-population tree prior \\
\verb;create_cep_tree_prior; & Create a coalescent exponential-population tree prior \\
\verb;create_yule_tree_prior; & Create a Yule tree prior \cite{yule} \\
\hline
\verb;create_beta_distr; & Create a beta distribution \\
\verb;create_exp_distr; & Create an exponential distribution \\
\verb;create_gamma_distr; & Create a gamma distribution \\
\verb;create_inv_gamma_distr; & Create an inverse gamma distribution \\
\verb;create_laplace_distr; & Create a Laplace distribution \\
\verb;create_log_normal_distr; & Create a log-normal distribution \\
\verb;create_normal_distr; & Create a normal distribution \\
\verb;create_one_div_x_distr; & Create a 1/X distribution \\
\verb;create_poisson_distr; & Create a Poisson distribution \\
\verb;create_uniform_distr; & Create a uniform distribution \\
\hline
\end{tabular}
\caption{beautier's functions}
\label{tab:functions}
\end{table}

In total, \verb;beautier; has 59 exported functions to create
a BEAST2 configuration file. \verb;beautier; is an alternative for a majority of BEAUti use cases.
\verb;beautier;s future extensions (also: road map) can be found on its GitHub.

\verb;beautier;, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

@@

\verb;beautier; has minimal support for parsing and interpreting BEAST2 output files,
for that the \verb;beastier; \cite{beastier} package is recommended.
